### 2023-02-04 proposal/proposal.qmd
Drop latex and go quarto

---
title: "Policy Project Proposal"
author: "Daniel Sánchez Pazmiño"
date: "Due Feb. 6th 2023"
bibliography: references.bib
style: apa.csl
format: pdf
knitr:
  opts_chunk: 
    echo: false
---

# Introduction

Ecuador is a middle income country located in upper South America next to Colombia and Peru. After its full dollarization in 2000, the country saw an economic surge between 2007-2014 as a result of high oil prices and a relatively indebted government [@Weisbrot.2017]. However, after a sharp decline in oil prices and the COVID-19 pandemic, the country seems to be unable to return to its pre-pandemic or pre-2014 levels, with employment indicators at an all-time low.

```{r}
#| label: employment-graph

# Load the data

employment<-
  write.csv('labour-market-stats.csv')

```

# Policy and institutional background

# Literature review