### 2023-02-04 proposal/proposal.qmd
commit

---
title: "The influence of legitimization of business on labour market formalization"
subtitle: "Policy Project Proposal"
author: "Daniel Sánchez Pazmiño"
date: "02/06/2023"
bibliography: references.bib
style: apa.csl
format: pdf
knitr:
  opts_chunk: 
    echo: false
    message: false
---

```{r}
#| label: setup

# Load libraries

library(tidyverse)

```

# Introduction

Ecuador is a middle income country located in upper South America next to Colombia and Peru. After its full dollarization in 2000, the country saw an economic surge between 2007-2014 as a result of high oil prices and a relatively indebted government [@Weisbrot.2017]. However, after a sharp decline in oil prices and the COVID-19 pandemic, the country seems to be unable to return to its pre-pandemic or pre-2014 levels, with employment indicators at an all-time low (see @fig-employment-graph).

```{r}
#| label: fig-employment-graph
#| fig-cap: "Adequate employment in Ecuador"

# Load the data

employment<-
  read.csv('labour-market-stats.csv')

# Graph

employment %>%
  filter(series == 'Adequate Employment') %>% 
  ggplot(aes(year, value))+
    geom_point(colour = '#2F66A9')+
    geom_line(colour = '#189ad3' )+
    theme_bw()+
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())+
    labs( x = 'Year',
          y = 'Adequate employment (% of labour force)')

```

# Policy and institutional background

# Literature review

# References